\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-09-23 introduction.tex
Adding a bib file

Let us provide an informal description of our results and setting.  For an
estimator $\theta_n$ with finite quadratic risk (mean squared error) $R_n =
\E_\theta[(\theta_n - \theta)^2]$, letting $\sigma^2 = \var(X)$
we are interested in particular in the
limit (assuming it exists)
\begin{equation}
  \label{eq:ARE_def}
  \lim_{n\to \infty} \frac{\sigma^2/n}{R_n},
\end{equation}
which describes the \emph{asymptotic relative efficiency} (ARE) of this
estimator compared to an estimator whose variance decreases as
$\sigma^2/n+o(n^{-1})$. Estimators of this form include the sample mean
and, under some conditions, the optimal Bayes estimator.

%% JCD: I AM HERE %%

In addition to the examples above, the excess risk in estimating a fixed parameter due to a one-bit per sample constraint is useful in bounding from below the excess risk in estimating signals from thier noisy one-bit quanitzed noisy measurements. Specifically, the excess risk or ARE we derive serve as the most optimistic estimate for the risk in estimating under one-bit per measurement constraint signals changing over time or space. Such estimation settings are considered in \cite{baraniuk2017exponential, jacques2013robust, plan2013one, li2017channel, choi2016near}. \par

@@

In setting (i), the estimator can evaluate the optimal mean estimator (e.g., the sample mean in the Gaussian case) and then quantize it using $n$ bits. Since the accuracy in describing the empirical mean decreases exponentially in $n$, the error due to quantization is negligible compared to the quadratic risk in estimating the mean \cite{720540}. Therefore, the ARE in this setting is $1$. Namely, asymptotically, there is no loss in performance due to the communication constraint under centralized encoding. 
%
In this paper we show that a similar result does not hold in setting (ii): the ARE of any adaptive estimation scheme is at most the ARE of the sample median. Specifically, when the samples are drawn from the normal distribution, this ARE equals $2/\pi$, showing that the one-bit constraint increases the effective sample size in estimating $\theta$ by at least $\pi/2$ compared to estimating it without the bit constraint. We also show that this lower bound on the ARE is tight by providing an estimator that attains it. In fact, we show that only one adaptiveness, as illustrated in Figure~\ref{fig:one_round} is enough to achieve the optimal ARE. 
%
Clearly, the penalty on the sample size in setting (iii) is at least as large as that in setting (ii). We show, however, that unlike in setting (ii), there is no distributed estimation scheme that is uniformly optimal in the sense that it attains the ARE of the sample median for all $\theta$ in the parameter space. This result is obtained by establishing local asymptotic normality of regularly enough encoding procedures for setting (iii), resulting in an exact characterization of the ARE of such procedures. To summarize our contributions, we establish that the optimal ARE in settings (i) is $1$, the optimal ARE in setting (ii) is ARE of the sample median, and no uniformly optimal procedure exists in setting (iii).

@@

It is important to note that although the ARE in setting (i) is one, this scheme already poses a non-trivial challenge for the design and analysis of optimal encoding and estimation procedures. Indeed, representing an unknown random quantity using $n$ bits is equivalent to designing a $2^n$ levels scalar quantizer \cite{gray1998quantization}. However, the optimal design of this quantizer depends on the distribution of its input, which is the goal of our estimation problem and hence its exact value is unknown. As a result, a non-trivial exploration-exploitation trade-off arises even in setting (i). 
%Note that the only missing parameter in our setting is the mean, which, under setting (i), is known to the encoder with uncertainty interval proportional to $\sigma/\sqrt{n}$. 
Therefore, while uncertainty due to quantization decreases exponentially in the number of bits $n$, hence the ARE is $1$, it is still difficult to derive an exact expression for the MSE in this setting. 
%
The situation is even more involved in the adaptive encoding setting (ii): an encoding and estimation strategy that is optimal for $n-1$ adaptive one-bit messages of a sample of size $n-1$ may not lead to a globally optimal strategy upon the recipient of the $n$th sample. Conversely, any one-step optimal strategy, in the sense that it finds the best one-bit message as a function of the current sample and the previous $n-1$ messages, is not guaranteed to be globally optimal. Therefore, while we characterize the optimal one-bit message given the previous messages, this characterization does not necessarily lead to an upper bound on the ARE. Instead, our result on the maximal ARE is obtained by bounding the Fisher information of any $n$ adaptive messages and using an appropriate information inequality. \par
%
%In addition to encoding and estimation schemes that lead to optimal results, we also consider two additional ``natural'' schemes. Specifically, in setting (ii) we consider the one-bit optimal scheme, i.e., the case of a greedy encoder that given the $n$th sample and the previous $n-1$ bits, provides a bit that minimizes the $n$-step MSE. In setting (iii) we also consider the case where the messages are obtained by comparing each sample against a prescribed threshold. This threshold may be different across samples and is assumed deterministic (independent of the data).

@@

When the variance $\sigma^2$ is negligible compared the to desired accuracy, the task of finding $\theta$ using one-bit queries in the adaptive setting (ii) is solved by a bisection style method over the parameter space. Therefore, the general case of non-zero variance is reminiscent of the noisy binary search problem with a possibly infinite number of unreliable tests \cite{cicalese2002least, Karp:2007:NBS:1283383.1283478}. However, since we assume a continuous parameter space, a more closely related problem is that of one-bit analog-to-digital conversion of a constant input corrupted by Gaussian noise. Using an SDM, Wong and Gray \cite{53738} showed that the output of the modulator converges to the true constant input almost surely, so that an SDM provides a consistent estimator for setting (ii). The rate of this convergence, however, was not analyzed and cannot be derived from the results of \cite{53738}. In particular, our results for setting (ii) imply that the asymptotic rate of convergence of the MSE in SDM to a constant input under an additive white Gaussian noise is at most $\sigma^2\pi/2$ over the number of feedback iterations. Baraniuk et. al \cite{baraniuk2017exponential} also considered adaptive one-bit measurements in the context of analog-to-digital conversion, although without noise at the input. By establishing the lower bound of $\sigma^2\pi/2n$ on the MSE, we show that the main results of \cite{baraniuk2017exponential}, an exponential MSE decaying rate, does not hold in the noisy setting. Stated otherwise, the MSE in the setting of \cite{baraniuk2017exponential} may decay exponentially up to the noise level, after which it decays at most as $\sigma^2\pi/2n$. \par

>>>>

Let us provide an informal description of our results and setting. For an
estimator $\theta_n$ with finite quadratic risk (mean squared error) $R_n =
\E_\theta[(\theta_n - \theta)^2]$, we study the limit
\begin{equation}
  \label{eq:ARE_def}
  \limsup_{n\to \infty} n R_n.
\end{equation}
By comparing this quantity to achievable rates of convergence without
communication constraints, we can evaluate the efficiency
losses---asymptotic relative efficiency---of the estimator to appropriately
optimal (unconstrained) estimators. (We shall be more formal in the sequel.)
By lower bounding the quantity~\eqref{eq:ARE_def}, we also provide limits on
estimation of single-bit-per-measurement constrained signals in more general
settings~\cite{baraniuk2017exponential, jacques2013robust, plan2013one,
  li2017channel, choi2016near}.

@@

In setting (i), the estimator can evaluate any optimal estimator of location
(e.g., the sample mean if the data is Gaussian), then quantize it using
$n$ bits. As the accuracy in describing the empirical mean decreases
exponentially in $n$, the quantization error is negligible compared
to the statistical error in mean estimation~\cite{720540}. That is,
centralized encoding induces no asymptotic efficiency loss.
%
The story is different in settings (ii) and (iii). Precisely, we show that
in the adaptive setting~(ii), the optimal efficiency of a one-bit scheme is
(asypmtotically) precisely that of the sample median, and that this
efficiency is achievable. As a concrete example, when $X_i$ are
i.i.d.\ Gaussian, we necessarily lose a factor of $\pi/2 \approx 1.57$ in
the asymptotic risk; the one-bit constraint decreases the effective sample
size by a factor of $\pi/2$ compared to estimating it without the bit
constraint. It turns out that, in the settings we consider,
only a \emph{single round} of adaptivity (see Fig.~\ref{fig:one_round} for
an illustration) is sufficient to achieve optimal convergence rates.
%
In distinction from setting (ii), in setting (iii) when the messages must be
independent, there is no distributed estimation scheme that achieves the
efficiency of the sample median uniformly over $\theta$.  We establish this
result via Le Cam's local asyptotic normality theory, allowing us
to provide exact characterizations of the asymptotic efficiency
of suitably regular encoding schemes.
%% To summarize our
%% contributions, we establish that the optimal ARE in settings (i) is $1$, the
%% optimal ARE in setting (ii) is ARE of the sample median, and no uniformly
%% optimal procedure exists in setting (iii).

@@

Our asymptotic setting is important in that it allows us to elide
difficulties present in finite sample settings. For example, in setting~(i),
developing an optimal quantizer at finite $n$ requires choosing a $2^n$
level scalar quantizer, which is non-trivial~\cite{gray1998quantization}.
In interactive and sequential settings (e.g.~(ii)), the situation is more
challenging, as it is unclear whether any type of compositionality applies,
in that an $n-1$-step optimal estimator may be only vaguely related to the
$n$-step optimal estimator. Thus, to provide our lower bounds, we rely on
stronger information-based inequalities, including the Van Trees
inequality~\cite{Tsybakov09} and Le Cam's local asymptotic normality
theory~\cite{LeCam86,LeCamYa00,VanDerVaart98}.

%% It is important to note that although the ARE in setting (i) is one, this
%% scheme already poses a non-trivial challenge for the design and analysis of
%% optimal encoding and estimation procedures. Indeed, representing an unknown
%% random quantity using $n$ bits is equivalent to designing a $2^n$ levels
%% scalar quantizer \cite{gray1998quantization}. However, the optimal design of
%% this quantizer depends on the distribution of its input, which is the goal
%% of our estimation problem and hence its exact value is unknown. As a result,
%% a non-trivial exploration-exploitation trade-off arises even in setting (i).
%% %Note that the only missing parameter in our setting is the mean, which, under setting (i), is known to the encoder with uncertainty interval proportional to $\sigma/\sqrt{n}$. 
%% Therefore, while uncertainty due to quantization decreases exponentially in
%% the number of bits $n$, hence the ARE is $1$, it is still difficult to
%% derive an exact expression for the MSE in this setting.
%
%% The situation is even more involved in the adaptive encoding setting (ii): an encoding and estimation strategy that is optimal for $n-1$ adaptive one-bit messages of a sample of size $n-1$ may not lead to a globally optimal strategy upon the recipient of the $n$th sample. Conversely, any one-step optimal strategy, in the sense that it finds the best one-bit message as a function of the current sample and the previous $n-1$ messages, is not guaranteed to be globally optimal. Therefore, while we characterize the optimal one-bit message given the previous messages, this characterization does not necessarily lead to an upper bound on the ARE. Instead, our result on the maximal ARE is obtained by bounding the Fisher information of any $n$ adaptive messages and using an appropriate information inequality. \par
%
%In addition to encoding and estimation schemes that lead to optimal results, we also consider two additional ``natural'' schemes. Specifically, in setting (ii) we consider the one-bit optimal scheme, i.e., the case of a greedy encoder that given the $n$th sample and the previous $n-1$ bits, provides a bit that minimizes the $n$-step MSE. In setting (iii) we also consider the case where the messages are obtained by comparing each sample against a prescribed threshold. This threshold may be different across samples and is assumed deterministic (independent of the data).

@@

When the variance $\sigma^2$ is negligible compared the to desired accuracy,
the task of finding $\theta$ using one-bit queries in the adaptive setting
(ii) is solved by a bisection style method over the parameter
space. Therefore, the general case of non-zero variance is reminiscent of
the noisy binary search problem with a possibly infinite number of
unreliable tests \cite{cicalese2002least,
  Karp:2007:NBS:1283383.1283478}. However, since we assume a continuous
parameter space, a more closely related problem is that of one-bit
analog-to-digital conversion of a constant input corrupted by Gaussian
noise. Using an SDM, Wong and Gray \cite{53738} showed that the output of
the modulator converges to the true constant input almost surely, so that an
SDM provides a consistent estimator for setting (ii). The rate of this
convergence, however, was not analyzed and cannot be derived from the
results of \cite{53738}. In particular, our results for setting (ii) imply
that the asymptotic rate of convergence of the MSE in SDM to a constant
input under an additive white Gaussian noise is at most $\sigma^2\pi/2$ over
the number of feedback iterations. Baraniuk et. al
\cite{baraniuk2017exponential} also considered adaptive one-bit measurements
in the context of analog-to-digital conversion, although without noise at
the input. By establishing the lower bound of $\sigma^2\pi/2n$ on the MSE,
we show that the main results of \cite{baraniuk2017exponential}, an
exponential MSE decaying rate, does not hold in the noisy setting. Stated
otherwise, the MSE in the setting of \cite{baraniuk2017exponential} may
decay exponentially up to the noise level, after which it decays at most as
$\sigma^2\pi/2n$. \par
